### 2022-09-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@f6666497809251353fa4cf0043de56f532cc817b 🚀

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Arsov \& Mirceva,
2019; Chami et al., 2022). Their main goal is to learn a low dimensional
vector representations for the nodes of the graph (embeddings), such
that key properties of the graph (e.g.~local or global structures) are
retained in the embedding space (Yan et al., 2005). Ecological networks
are an interesting candidate for the widespread application of
embeddings, as they tend to possess a shared structural backbone (Bramon
Mora et al., 2018), which hints at structural invariants that can be
revealed at lower dimensions. Indeed, food webs are inherently
low-dimensional objects, and can be adequately represented with less
than ten dimensions (Braga et al., 2019; Eklöf et al., 2013). Simulation
results by Botella et al. (2022) suggest that there is no best method to
identify architectural similarities between networks, and that multiple
approaches need to be tested and compared to the network descriptor of
interest. This matches previous, more general results on graph
embedding, which suggest that the choice of embedding algorithm matters
for the results (Goyal \& Ferrara, 2018). In tbl.~\ref{tbl:methods}, we
present a selection of common graph and node embedding methods,
alongside examples of their use to predict species interactions; most of
these methods rely either on linear algebra, or on pseudo-random walks
on graphs.

>>>>

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Arsov \& Mirceva,
2019; Chami et al., 2022). Their main goal is to learn a low dimensional
vector representations for the nodes of the graph (embeddings), such
that key properties of the graph (e.g.~local or global structures) are
retained in the embedding space (Yan et al., 2005). Ecological networks
are an interesting candidate for the widespread application of
embeddings, as they tend to possess a shared structural backbone (see
\emph{e.g.} Bramon Mora et al., 2018 for food webs), which hints at
structural invariants in empirical data; assuming that these structural
invariants are indeed widespread, they would dominate the structure of
networks, and therefore be adequately captured by the first (lower)
dimensions of an embedding, without the need to measure derived aspects
of their structure (\emph{e.g.} motifs, paths, modularity, \ldots).

Indeed, food webs are inherently low-dimensional objects, and can be
adequately represented with less than ten dimensions (Braga et al.,
2019; Eklöf et al., 2013). Simulation results by Botella et al. (2022)
suggest that there is no best method to identify architectural
similarities between networks, and that multiple approaches need to be
tested and compared to the network descriptor of interest. This matches
previous, more general results on graph embedding, which suggest that
different embedding algorithms yield different network embeddings (Goyal
\& Ferrara, 2018), calling for a careful selection of the
problem-specific approach to use. In tbl.~\ref{tbl:methods}, we present
a selection of common graph and node embedding methods, alongside
examples of their use to predict species interactions; most of these
methods rely either on linear algebra, or on pseudo-random walks on
graphs.
